Multi-level modeling with Openflexo/\FML - A contribution to the \mpc

### 2021-05-31 6-modelpresentation.tex
Updates from Overleaf

\texttt{ProcessType} is a specialization of the \texttt{Type} concept, and references a collection of \texttt{TaskType} through the composition relation \texttt{taskTypes} with (0..*) cardinality (\textbf{P1}). A \texttt{TaskType} is embedded in a \texttt{ProcessType} and inherits from its context. To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \textsf{Claim Handling}, instance of \texttt{ProcessType}, and \textsf{Receive Claim}, \textsf{Assess Claim} and \textsf{Pay premium}, instances of \texttt{TaskType}.

@@

The same pattern applies for the used and produced artifacts by roles \texttt{usedArtifacts}, \texttt{producedArtifacts} and \texttt{performingActors} defined in \texttt{Task} (\textbf{P13}). An instance of \texttt{Artifact} specializes \textit{MultiInstance} and references a set of \texttt{ArtifactType} through the specialized \texttt{/types} role (\textbf{P14} and \textbf{P16}). Likewise, \texttt{Actor} specializes \texttt{MultiInstance} and references a set of \texttt{ActorType} through the specialized \texttt{/types} relation (\textbf{P15}).

The semantics is unclear relatively to the instantiation policy for artifacts. In the requirements, nothing is explicitly stated about the link between the \emph{artifact type} of a \emph{task type} and the type of the corresponding artifact

We assume that task execution implies that for each used and produced \emph{artifact type} defined in a related \emph{task type}, it exists at least one \emph{artifact} declaring required ontologic instantiation (through \textit{type} role). The following excerpt of \FML code shows a partial implementation of this. The same pattern applies to produced artifacts.
\noteFabien{Je ne suis pas bien sûr de comprendre le paragraphe}

@@

\textsf{'Requirements analysis'} is an instance of \texttt{SETaskType}, \textsf{'Analyst'} is an instance of \texttt{ActorType} and \textsf{'Requirement specifications'} is an instance of \texttt{SEArtifactType}. The value of the property \texttt{producedArtifactTypes} of \textsf{'Requirements analysis'} is \texttt{\{}\textsf{Requirement specifications}\texttt{\}} and is \texttt{\{}\textsf{Analyst}\texttt{\}} for \texttt{allowedActorTypes} (\textbf{S1}). Similarly, \textsf{'Test case design'} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \textsf{'Analyst'} by the \texttt{allowedActorTypes} relation. \textsf{'Test case design'} produces \textsf{'Test cases'} instance of \texttt{SETaskType}. The latter is used in \textsf{'Test design review'} (\texttt{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).

@@

\textsf{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:

@@

The requirement \textbf{S8} is simply expressed by the definition of the \textsf{'Testing'} instance of \texttt{SETaskType}, \textsf{'Tester'} instance of \texttt{ActorType}, and \textsf{'Test report'} instance of \texttt{ArtifactType}.

>>>>

\texttt{ProcessType} is a specialization of the \texttt{Type} concept, and references a collection of \texttt{TaskType} through the composition relation \texttt{taskTypes} with (0..*) cardinality (\textbf{P1}). A \texttt{TaskType} is embedded in a \texttt{ProcessType} and inherits from its context. To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \inst{Claim Handling}, instance of \texttt{ProcessType}, and \inst{Receive Claim}, \inst{Assess Claim} and \inst{Pay premium}, instances of \texttt{TaskType}.

@@

The same pattern applies for the used and produced artifacts by roles \texttt{usedArtifacts}, \texttt{producedArtifacts} and \texttt{performingActors} defined in \texttt{Task} (\textbf{P13}). An instance of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of \texttt{ArtifactType} through the specialized \texttt{/types} role (\textbf{P14} and \textbf{P16}). Likewise, \texttt{Actor} specializes \texttt{MultiInstance} and references a set of \texttt{ActorType} through the specialized \texttt{/types} relation (\textbf{P15}).

The semantics is unclear relatively to the instantiation policy for artifacts. In the requirements, nothing is explicitly stated about the link between the \emph{artifact type} of a \emph{task type} and the types of the artifacts of a corresponding task.
Here, we assume that task execution implies that for each used and produced \emph{artifact type} defined in a related \emph{task type}, it exists at least one artifact of the right type. The following excerpt of \FML code shows a partial implementation of this. The same pattern applies to produced artifacts.

@@

\inst{Requirements analysis} is an instance of \texttt{SETaskType}, \inst{Analyst} is an instance of \texttt{ActorType} and \inst{Requirement specifications} is an instance of \texttt{SEArtifactType}. The value of the property \texttt{producedArtifactTypes} of \inst{Requirements analysis} is \texttt{\{}\inst{Requirement specifications}\texttt{\}} and is \texttt{\{}\inst{Analyst}\texttt{\}} for \texttt{allowedActorTypes} (\textbf{S1}). Similarly, \inst{Test case design} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \inst{Analyst} by the \texttt{allowedActorTypes} relation. \inst{Test case design} produces \inst{Test cases} instance of \texttt{SETaskType}. The latter is used in \inst{Test design review} (\texttt{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).

@@

\inst{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:

@@

The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}.
